How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-26 tse-paper/content.tex
Include review until page 2

In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that automatic software build
processes create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over time~\cite{}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information in it which is linked to the error, is akin to
searching for a needle in a haystack.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,machalica2019predictive},
or investigate static analysis tools in CI
builds~\cite{zampetti2017open}.

@@

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature study (step 1).
The literature
survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
natural language processing, information retrieval, or searching
for keywords.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
% C: 14 is still quite big, it's 'impl available' when we saw
% no overlap.
Decide whether you want to keep this in :)
only 14 independent implementations are available, one of which stems
from the second author of this paper.
Out of these implementations, all are based on either a custom parser
or regular expressions.
%Using 1) a set of
% manually curated regular expressions,
% similar to TravisTorrent~\cite{beller2017oops},
% 2) information
% retrieval techniques,
% % Can you shortly explain the technique here?
% and 3) a list of search keywords
% to identify regions of interest in the build log.

@@

% TODO caro: maybe convert this into a proper ordered list?
As a consequence, we took it upon ourselves (step 2) to provide
implementations of the three most promising automatic build log analysis
techniques we observed in the literature survey:
a technique 1) able to automatically synthesize regular expressions,
called \emph{Program Synthesis by Example (PBE)}, 2) based on information
retrieval, using \emph{Common Text Similarity (CTS)}  and 3)
inspired by the manual adhoc-search for key parts in the log, employing
\emph{Keyword Search (KWS)}.
Even though these three approaches---regular expressions, text similarity,
keyword search---
have different strengths and weaknesses, the literature
survey shows that their use often seems arbitrary.
As a consequence,
in the second research question we ask:

\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
2}}
How do build log analysis techniques compare against each other?
\end{simplebox}

To answer this question, as a third step,
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
However,
% TODO 1-2 sentences about results, inclduing some accuracy measures

Finally, due to this ambiguity, we develop a set of guidelines
(\Cref{fig:overview}, step 4)
on how to choose the most suitable build log analysis
technique for the task at hand.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid-solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

@@

In this section, we survey the emerging field of build log
analysis.
We start with a differentiation to the related field of
system log analysis.
% ref to fig 2 so it can be here and above fig 3
Then, we describe the methodology (\Cref{fig:lit-survey})
and results of a systematic mapping
study to
determine the state-of-research on build log analysis following the
guidelines by Petersen et
al.~\cite{petersen2008systematic,petersen2015guidelines}.
We close by summarizing how build logs are commonly
analyzed in literature and discuss the implications of these findings.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference here is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

@@

\begin{simplebox}[minipage boxed title*=-1.5cm,
attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ2:} How do build log analysis techniques compare against
each other?}
\begin{itemize}[leftmargin=1.2cm]
  \item[\textbf{RQ2.1:}] How many examples does a technique need to
  perform best?
  \item[\textbf{RQ2.2:}] How structurally similar do the examples
  need for a technique to be applicable?
  \item[\textbf{RQ2.3:}] How accurate are the retrievals of a technique?
\end{itemize}
\end{simplebox}

>>>>

In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that automatic software build
processes create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over time~\cite{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information in it which is linked to the actual error, is akin to
searching for a needle in a haystack. The large numbers of unrelated, but amibguous warnings that many build logs regularly contain, further compounds this problem.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,machalica2019predictive},
or investigate the role of automated static analysis tools (ASATs, \cite{beller2016analyzing}) in CI
builds~\cite{zampetti2017open}.

@@

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature study (step 1).
The literature
survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
searching for keywords, or performing natural language processing.
However, there is currently no guidance on when to use which of these approaches.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
the few available implementations are all based on either a custom parser
or regular expressions. These two approaches in particular are problematic, as their creation and maintenane is manual and expensive, does not generalize across build logs from different build environments, and is, as the experience of TravisTorrent has shown, prone to error. In fact, some even ``strongly discourage'' the creation of such
bespoke parsers~\cite{urli2018design}. In summary, the outcome of our literature survey shows the need for a new
generation of automated techniques to analyze build logs, which are documented in detail and which are openly available, together with guidance on when to use which tool.

%Using 1) a set of
% manually curated regular expressions,
% similar to TravisTorrent~\cite{beller2017oops},
% 2) information
% retrieval techniques,
% % Can you shortly explain the technique here?
% and 3) a list of search keywords
% to identify regions of interest in the build log.

@@

% TODO caro: maybe convert this into a proper ordered list?
To fill this gap, we adapted and implemented three novel, prototypical implementations to
automatically analyze build logs. The ideas behind the implementations are based on promising approaches adopted from other research fields and automatic build log analysis
techniques:
a technique 1) able to automatically synthesize regular expressions,
called \emph{Program Synthesis by Example (PBE)}, 2) based on information
retrieval, using \emph{Common Text Similarity (CTS)}  and 3)
inspired by the manual adhoc-search for key parts in the log, employing
\emph{Keyword Search (KWS)}.
These three approaches---regular expressions, text similarity,
keyword search---
have different strengths and weaknesses.
As a consequence,
in the second research question we ask:

\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
2}}
How do build log analysis techniques compare?
\end{simplebox}

To answer this question, as a third step,
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
However,
% TODO Caro: 1-2 sentences about results, inclduing some accuracy measures

Finally, due to this ambiguity, we develop a set of guidelines
(\Cref{fig:overview}, step 4)
on how to choose the most suitable build log analysis
technique for the task at hand.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

@@

In this section, we survey the emerging field of build log
analysis.
We start with a differentiation to the field of
system log analysis.
% ref to fig 2 so it can be here and above fig 3
Then, we describe the methodology (\Cref{fig:lit-survey})
and results of a systematic mapping
study to
determine the state-of-research on build log analysis following the
guidelines by Petersen et
al.~\cite{petersen2008systematic,petersen2015guidelines}.
We close by summarizing how build logs are commonly
analyzed in literature and discuss the implications of these findings.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

@@

\begin{simplebox}[minipage boxed title*=-1.5cm,
attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ2:} How do build log analysis techniques compare?}
\begin{itemize}[leftmargin=1.2cm]
  \item[\textbf{RQ2.1:}] How many examples does a technique need to
  perform best?
  \item[\textbf{RQ2.2:}] How structurally similar do the examples
  need for a technique to be applicable?
  \item[\textbf{RQ2.3:}] How accurate are the retrievals of a technique?
\end{itemize}
\end{simplebox}
